%!TEX root = ../thesis.tex
% ******************************* Thesis Appendix A ****************************
\chapter{}

\section{Benchmarking functions}

The following is a list of synthetic functions and real datasets presented in previous works, where the goal is to evaluate bayesian optimization algorithms. 

\begin{enumerate}
\item \citep{Garnett2013} Synthetic in-model data matching the proposed model, with $d=2, 3$, and $D=10, 20$.
\item \citep{Garnett2013} (Synthetic) Braning function, $d=2$, hidden in a higher dimensional space $D=10, 20$.
\item \citep{Garnett2013} Temperature data $D=106$ and $d=2$.
\item \citep{Garnett2013} Communities and Crime dataset $d=2$, and $D=96$. 
\item \citep{Garnett2013} Relative location of CT slices on axial axis with $d = 2$ and $D=318$. 
\item \citep{Djolonga2013} (Synthetic) Random GP samples from 2-dimensional Matern-Kernel-output, embedded within 100 dimensions
\item \citep{Djolonga2013} Gabor Filters: Determine visual stimuli that maximally excite some neurons which reacts to edges in the image.
We have $f(x) = \exp( -( \theta^T x - 1 )^2 )$. $\theta$ is of size 17x17, and the set of admissible signals is $d$.
\item \citep{Wang2013} (Synthethic) $d=2$ and $D=1*10^9$.
\item \citep{Wang2013} $D=47$ where each dimension is a parameter of a mixed integer linear programming solver.
\item \citep{Wang2013} $D=14$ with $d$ for a random forest body part classifier.
\item \citep{Tripathy}  (Synthetic) Use $d=1,10$ and $D=10$.
\item \citep{Tripathy} (Half-synthetic) Stochastic elliptic partial differential equation, where $D=100$, and an assume value for $d$ of $1$ or $2$.
\item \citep{Tripathy} Granular crystals $X \in \mathbb{R}^{1000 \times 2n_p +1}$, and $y \in \mathbb{R}^{1000}$.
\item \citep{Gardner2017} (Synthetic) Styblinski–Tang function where $D$ is freely choosable.
\item \citep{Gardner2017} (Synthetic) Michalewicz function where $D$ is freely choosable.
\item \citep{Gardner2017} (Simulated) NASA cosmological constant data where $D=9$.
\item \citep{Gardner2017} Simple matrix completion with $D=3$.
\item \citep{Rana2017} (Synthetic) Hertmann6d in [0, 1].
\item \citep{Rana2017} (Synthetic) Unnormalized Gaussian PDF with a maximum of 1 in $[-1, 1]^d$ for $D=20$ and $[-0.5, 0.5]^d$ for $D=50$
\item \citep{Rana2017} (Synthetic) Generalized Rosenbrock function $[-5, 10]^d$
\item \citep{Rana2017} Training cascade classifiers, with $D=10$ per group.
\item \citep{Rana2017} Optimizing alloys $D=13$. 
\item \citep{Li2018} (Synthetic) Gaussian mixture function
\item \citep{Li2018} (Synthetic) Schwefel's 1.2 function.
\item \citep{OptimizationTestFunctions} A list of optimiztation test functions can be found here.
\item \citep{Jamil2013} A more comprehensive list of general functions can be found here.
%\item https://github.com/automl/HPOlib2/tree/master/hpolib/benchmarks/synthetic_functions
\end{enumerate}